---
title: "Adaptive Sampling using R"
author: "Aaron Shaffer"
date: "May 3, 2017"
output: slidy_presentation
---
## Adaptive Sampling: (Refresher)
<style>
.column-left{
  float: left;
  width: 50%;
  text-align: left;}
.column-center{
  display: inline-block;
  width: 100%;
  text-align: center;
}
.column-right{
  float: right;
  width: 50%;
  text-align: left;
}

/* 5 columns */
.col-xs-15,
.col-sm-15,
.col-md-15,
.col-lg-15 {
    position: relative;
    min-height: 1px;
    padding-right: 10px;
    padding-left: 10px;
}

.col-xs-15 {
    width: 20%;
    float: left;
}
@media (min-width: 768px) {
    .col-sm-15 {
        width: 20%;
        float: left;
    }
}
@media (min-width: 992px) {
    .col-md-15 {
        width: 20%;
        float: left;
    }
}
@media (min-width: 1200px) {
    .col-lg-15 {
        width: 20%;
        float: left;
    }
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
pkgs <-c('reshape2','dplyr','ggplot2','plotly','igraph','plotrix')
for(p in pkgs)
  if(p %in% rownames(installed.packages()) == FALSE) {install.packages(p)}
for(p in pkgs)
  suppressPackageStartupMessages(library(p, quietly=TRUE, character.only=TRUE))
rm('p','pkgs')

melt.matrix <- function(world){
  suppressMessages(world %>% 
  data.frame %>% 
  melt() %>% 
  mutate(id = seq_len(nrow(world) * ncol(world)), 
        x = rep(c(1:ncol(world)),nrow(world)),
        y = sort(rep(c(1:nrow(world)),ncol(world)))) %>%
    data.frame() %>% select(value,id,x,y)
  )
}

getid <- function(x,y,ncol){
  return(x + ((y-1) * ncol))
}

plotly.adjmatrix <- function(world, name){
  gg <- melt.matrix(world) %>% 
    ggplot(aes(x = as.factor(x), 
               y = as.factor(y),
               "X" = x,
               "Y" = y,
               "ID" = id)) + 
    geom_raster(aes(fill = value, alpha = value/max(value))) + 
    scale_fill_gradient(low="grey90",high="red") + 
    geom_text(aes(label = ifelse(value,value,""))) +
    theme(panel.background = element_blank(), legend.position = "none") + 
    labs(x = "X",y = "Y") + ggtitle(name)
  
  ggplotly(gg, tooltip = c("X","Y","value","id"))
}

plotly.adjmatrix.id <- function(world, name){
  gg <- melt.matrix(world) %>% 
    ggplot(aes(x = as.factor(x), 
               y = as.factor(y),
               "value" = value,
               "X" = x,
               "Y" = y,
               "ID" = id)) + 
    geom_tile(aes(fill = as.factor(id)),color = "black") + 
    geom_text(aes(label = id)) +
    theme(panel.background = element_blank(), legend.position = "none") + 
    labs(x = "X",y = "Y") + ggtitle(paste("Index values of ",name))
  
  ggplotly(gg, tooltip = c("X","Y","value","id"))
}

```

"10.7: Adaptive Sampling: Suppose that we had a rectangular forest for which you wish to estimate the number of trees of a certain
species. One way to accomplish this is to lay a grid over the forest a sample a certain number of cells. But suppose that these trees tend to occur in clusters so that if you see one tree in a cell then you are likely to see other trees in nearby cells. Why not visit these neighboring cells while in the other? This is an example of adaptive sampling. Instead of a sample size being selected in advance and being fixed the sample size is allowed to vary as new information is coming in."

<div class="column-left">

An unbiased estimator of the mean density per cell in the population is given by:

$\hat{\mu} = \frac{1}{n}\sum\limits_{i=1}^{n}{\frac{y_i}{m_i}} = \frac{1}{n}\sum\limits_{i=1}^{n}{\bar{y_i}}$

Which is the means of the network means.  The estimated variance of this estimator is what you would expect from simple random sampling.

$\hat{V}(\hat{\mu}) = (1 - \frac{n}{N}) \frac{s_\bar{y}^2}{n}$

Where: $\frac{s_\bar{y}^2}{n}$

is the sample variance of the network means

ie: $s_\bar{y}^2 = var(\frac{y_i}{m_i})$

</div>

<div class ="column-right">

In R:

```{r, eval = FALSE, echo = TRUE}
mu.hat <- 1/n * sum(y/m)

s.ybar.2 <- var(y/m)

V.hat.mu.hat <- (1 - n/N) * s.ybar.2/n
```

</div>

## Example from our notes:

<div class="column-left">
$n = 5$

$N = 100$

|     |   |   |   |   |   |
|:---:|:-:|:-:|:-:|:-:|:-:|
|$m_i$| 1 | 4 | 1 | 5 | 2 |
|$y_i$| 0 | 12 | 1 | 15 | 7 |

$\hat{\mu} = \frac{1}{5}\sum{(\frac{0}{1} + \frac{12}{4} + \frac{1}{1} + \frac{15}{5} + \frac{7}{2})} = 2.1$

$s_\bar{y}^2 = var(\frac{y}{m}) = 2.3$

$\hat{V}(\hat{\mu}) = (1 - \frac{5}{100}) \frac{2.3}{5} = 0.437$

</div>
<div class="column-right">
```{r, echo = TRUE}
n <- 5
N <- 100
m <- c(1,4,1,5,2)
y <- c(0,12,1,15,7)
```

```{r, echo = TRUE}
adaptive.sample.confint.fun <- function(y,m,n,N) {
  mu.hat <- 1/n * sum(y/m) 
  s.y.2 <- var(y/m)
  v.mu <- (1 - n/N) * s.y.2/n
  c("mu.hat" = mu.hat, 
    "s.y.2" = s.y.2,
    "v.mu" = v.mu,
    "var" = 2 * sqrt(v.mu),
    "lower" = mu.hat - 2 * sqrt(v.mu),
    "upper" = mu.hat + 2 * sqrt(v.mu))
}

adaptive.sample.confint.fun(y,m,n,N)
```

</div>
<div class="column-center">

So for this sample we estimate that the true mean density of our network was between 0.77 and 3.42 Maple Trees per cell 

</div>

## What might this world have originally looked like?

Data Structure: 2d Matrix

in R: ?matrix

Hint: "lay a grid over the forest"

<div class="column-left">
```{r, echo = TRUE}
example.world <- matrix(c( 1, 0, 3, 3, 0, 0,-1,
                           0, 0, 3, 0, 3, 0,-1,
                           3, 0, 0, 0, 0, 0, 0,
                           0, 4, 0, 0, 0, 3, 3,
                           0, 0, 0, 3, 3, 3, 0,
                          -1,-1, 0, 0, 0, 0, 0, 
                          -1,-1,-1,-1,-1,-1,-1),
                        nrow = 7, ncol = 7,byrow=TRUE)
example.world
```

If you want to rebuild a world given some adaptive samples:

Think of a world like a game of minesweeper.

Since we know the y and m for a few samples we know a couple of things.  Anything around these samples has to have a value of 0, as defined by what it means to take an adaptive sample.

Any square in the representation of the world that doesnt touch a known sample has some unknown value "-1"

</div>
<div class="column-right">
```{r}
plotly.adjmatrix(example.world,"example.world, -1 means you cannot infer what value might be there")
```
</div>

## How do we navigate this world on the computer?

<div class="column-left">

First lets look at another example world:

```{r, echo = TRUE}
figure10.2 <- matrix(rep(rep(0,10),10),
  nrow = 10, ncol = 10)

figure10.2[,10] <- c(4,0,0,0,0,0,0,0,0,7)
figure10.2[,9]  <- c(3,0,2,1,0,0,0,0,4,1)
figure10.2[,8]  <- c(0,0,0,4,5,0,0,3,2,0)
figure10.2[,7]  <- c(5,4,0,0,0,0,0,0,0,0)
figure10.2[,6]  <- c(0,0,0,2,0,1,0,0,7,0)
figure10.2[,5]  <- c(0,5,0,4,0,0,0,0,0,0)
figure10.2[,4]  <- c(0,0,3,0,0,4,0,3,0,0)
figure10.2[,3]  <- c(0,1,0,0,0,3,0,0,4,0)
figure10.2[,2]  <- c(0,0,0,2,6,0,0,0,0,0)
figure10.2[,1]  <- c(0,4,0,4,0,0,2,3,0,0)
```

This world is represented as a Matrix in memory.

In order to navigate this world and take an adpative sample this data structure will have to be reshaped in order to analyze its contents.

To understand the analysis some definitions of functions and parts of the algorithm used to take the adaptive sample need to be known.

</div>

<div class="column-right">

```{r}
plotly.adjmatrix(figure10.2,"Figure 10.2")
```
</div>

## Definition 1: "Adjacency List"

"An adjacency list is a collection of unordered lists used to represent a finite graph.  Each list represents the set of neighbors of a vertex of a graph"

Step 1. Represent the matrix as a graph where each cell of the matrix is a vertex on a graph.

Three functions created 

Adjlist function:

  - Given a world as a matrix object:
        
    - Return the full adjacency list for every node of that matrix as if it were a graph
        
Adjlist.adaptive functiion:

  - Given a world as a matrix object
    
    - Return the adjacency list for a world that obeys the rules of an adaptive sampling grid

      - ie: In order for a vertex to have any other verticies in its adjacency list it must have a non zero value
        
getEdges function:

  - Given a world as a matrix object as well as a boolean for Adaptive Or Not
    
      - Returns an adjacency list that the igraph package can use inorder to plot a graph

## How to generate the Adjlist:

<div class="column-left">
```{}
for each index "i" of the matrix:
    - Is "i" in the Top Row?
    - Is "i" in the Bottom Row?
    - Is "i" in the Leftmost column?
    - Is "i" in the Rightmost Column?

    if (Not in Leftmost column)
      left <- i - 1
    
    if (Not in the Leftmost column or Topmost row)
      upleft <- i - 1 + ncol
    
    if (Not in the Topmost Row)
      up <- i + ncol
    
    if (Not in the Topmost row or Rightmost column)
      upright <- i + 1 + ncol
    
    if (Not in the Rightmost column)
      right <- i + 1
    
    if (Not in the Bottommost row or the Rightmost column)
      downright <- i + 1 - ncol
    
    if (Not in the Bottommost row)
      down <- i - ncol
```
</div>

<div class="column-right">

all graphs use Plotly mouse over a vertex in the graph for more information

```{r}
plotly.adjmatrix.id(figure10.2, "Figure 10.2")
```
</div>

## Adjlist Functions:

<div class="column-left">
```{r, echo = TRUE}
adjlist <- function(world){
  nrow <- nrow(world)
  ncol <- ncol(world)
  
  G <- as.vector(world)
  
  adj.list <- rep(list(c()),nrow*ncol)  
  for(index in 1:length(G)){
    last.column <- index %% ncol == 0
    first.column <- (index - 1) %% ncol == 0
    first.row <- index <= ncol
    last.row <- index > (nrow * ncol - ncol)
    
    left <- ifelse(first.column,NA,index - 1)
    up.left <- ifelse(last.row | first.column,NA,index + ncol - 1)
    up <- ifelse(last.row,NA,index + ncol)
    up.right <- ifelse(last.row | last.column, NA,index + ncol + 1)
    right <- ifelse(last.column, NA, index + 1)
    down.right <- ifelse(first.row | last.column, NA, index - ncol + 1)
    down <- ifelse(first.row, NA, index - ncol)
    down.left <- ifelse(first.row | first.column , NA, index - ncol - 1)
    
    adjlist.i <- c()
    if(!is.na(left))
      adjlist.i <- c(adjlist.i,"W" = left)
    if(!is.na(up.left))
      adjlist.i <- c(adjlist.i,"NW" = up.left)
    if(!is.na(up))
      adjlist.i <- c(adjlist.i,"N" = up)
    if(!is.na(up.right))
      adjlist.i <- c(adjlist.i,"NE" = up.right)
    if(!is.na(right))
      adjlist.i <- c(adjlist.i,"E"= right)
    if(!is.na(down.left))
      adjlist.i <- c(adjlist.i,"SE" = down.left)
    if(!is.na(down))
      adjlist.i <- c(adjlist.i,"S" = down)
    if(!is.na(down.right))
      adjlist.i <- c(adjlist.i,"SW" = down.right)
    
    adj.list[[index]] <- adjlist.i
  }
  return(adj.list)
}
```
</div>

<div class="column-right">
```{r, echo = TRUE}
adjlist.adaptive <- function(world){
  nrow <- nrow(world)
  ncol <- ncol(world)
  
  G <- as.vector(world)
  
  adj.list <- rep(list(c()),nrow*ncol)  
  for(index in 1:length(G)){
    if(G[index] != 0) {
      last.column <- index %% ncol == 0
      first.column <- (index - 1) %% ncol == 0
      first.row <- index <= ncol
      last.row <- index > (nrow * ncol - ncol)
      
      left <- ifelse(first.column,NA,index - 1)
      up.left <- ifelse(last.row | first.column,NA,index + ncol - 1)
      up <- ifelse(last.row,NA,index + ncol)
      up.right <- ifelse(last.row | last.column, NA,index + ncol + 1)
      right <- ifelse(last.column, NA, index + 1)
      down.right <- ifelse(first.row | last.column, NA, index - ncol + 1)
      down <- ifelse(first.row, NA, index - ncol)
      down.left <- ifelse(first.row | first.column , NA, index - ncol - 1)
      
      to.return <- c()
      if(!is.na(left)) {
        if(G[left] != 0) {
          to.return <- c(to.return,"W" = left) 
        }
      }
      if(!is.na(up.left)){
        if(G[up.left] != 0) {
          to.return <- c(to.return,"NW" = up.left)  
        }
      }
      if(!is.na(up)) {
        if(G[up] != 0){
          to.return <- c(to.return,"N" = up)
        }
      }
      if(!is.na(up.right)) {
        if(G[up.right] != 0){
          to.return <- c(to.return,"NE" = up.right)
        }
      }
      if(!is.na(right)) {
        if(G[right] != 0){
          to.return <- c(to.return,"E"= right)
        }
      }
      if(!is.na(down.left)) {
        if(G[down.left] != 0){
          to.return <- c(to.return,"SE" = down.left)
        }
      }
      if(!is.na(down)) {
        if(G[down] != 0){
          to.return <- c(to.return,"S" = down)
        }
      }
      if(!is.na(down.right)) {
        if(G[down.right] != 0){
          to.return <- c(to.return,"SW" = down.right)
        }
      }
      
      adj.list[[index]] <- to.return
    } else {
      adj.list[[index]] <- c("NA" = NA)
    }
  }
  return(adj.list)
}
```
</div>

## Non adaptive figure 10.2

<div class="column-left">
```{r}
plotly.adjmatrix.id(figure10.2, "Figure 10.2")
```

```{r}
figure10.2.adjlist <- adjlist(figure10.2)
```

A matrix where each cell is the vertex of a graph and all vertecies of this graph are fully connected to each of their neighbors

ie: Any vertex to the NE, N, NW, W, SW, S, SE, E is connected

</div>

```{r}
getEdges <- function(world,adaptive){
  if(adaptive == TRUE)
    world.adj <- adjlist.adaptive(figure10.2)
  else
    world.adj <- adjlist(figure10.2)
  edge.list <- c()
  for(i in 1:length(world.adj)){
    for(j in 1:length(world.adj[[i]])){
      if(i <= ifelse(!gtools::invalid(world.adj[[i]][j]),world.adj[[i]][j],0)){
        edge.list <- c(edge.list,i,ifelse(!gtools::invalid(world.adj[[i]][j]),world.adj[[i]][j],i))
      }
    }
  }
  edge.list
}
```

<div class="column-right">

```{r}
g1 <- graph(edges = rev(getEdges(figure10.2,FALSE)), 
            directed = F, n = 100)
plot(g1, layout=layout_on_grid)
```

The edges of this world

</div>


## Adj list for a function following the rules of adaptive sampling

<div class ="column-left">

```{r}
plotly.adjmatrix(figure10.2, "Figure 10.2")
```

```{r}
figure10.2.adj <- adjlist.adaptive(figure10.2)
```

</div>
<div class="column-right">

```{r}
g2 <- graph(edges = rev(getEdges(figure10.2,TRUE)), 
            directed = F, n = 100)
plot(g2, layout=layout_on_grid)
```

What the world looks like when edges are only allowed between vertices with non 0 values

</div>

## Compare the differences in the adjacency lists

<div class="column-left">
```{r}
g1 <- graph(edges = rev(getEdges(figure10.2,FALSE)), 
            directed = F, n = 100)
plot(g1, layout=layout_on_grid)
```

```{r}
plotly.adjmatrix.id(figure10.2,"Figure 10.2")
```

Full adjaceny list:
```{r, echo = TRUE}
adjlist(figure10.2)
```

</div>

<div class="column-right">
```{r}
g2 <- graph(edges = rev(getEdges(figure10.2,TRUE)), 
            directed = F, n = 100)
plot(g2, layout=layout_on_grid)
```

```{r}
plotly.adjmatrix(figure10.2,"Figure 10.2")
```

Full Adaptive sampling Adjacency list:
```{r, echo = TRUE}
adjlist.adaptive(figure10.2)
```
</div>

## Definition 2: "Queue"

"In computer science, a queue is a particular kind of abstract data type or collection in which the entities in the collection are kept in order and the principal (or only) operations on the collection are the addition of entities to the rear terminal position, known as enqueue, and removal of entities from the front terminal position, known as dequeue. This makes the queue a First-In-First-Out (FIFO) data structure. In a FIFO data structure, the first element added to the queue will be the first one to be removed." -- Wikipedia

```{}
## Queue$Push() <- "Enqueue" push to the back of the Queue
## Queue$Pop() <- "Dequeue", remove the first element in the Queue
## Queue$Empty() <- returns Queue$size() == 0
```

```{r}
Queue <- setRefClass(Class = "Queue",
                     fields = list(
                       name = "character",
                       data = "list"
                     ),
                     methods = list(
                       size = function() {
                         'Returns the number of items in the queue.'
                         return(length(data))
                       },
                       #
                       push = function(item) {
                         'Inserts element at back of the queue.'
                         data[[size()+1]] <<- item
                       },
                       #
                       pop = function() {
                         'Removes and returns head of queue (or raises error if queue is empty).'
                         if (size() == 0) stop("queue is empty!")
                         value <- data[[1]]
                         data[[1]] <<- NULL
                         value
                       },
                       #
                       initialize=function(...) {
                         callSuper(...)
                         #
                         # Initialise fields here (place holder)...
                         #
                         .self
                       },
                       empty=function(...){
                         return(size() == 0)
                       }
                     )
)
```

Adapted from https://github.com/DataWookie/liqueueR/blob/master/R/queue.R

## How is a sample taken?

From the definition of an adaptive sample:

"..But suppose that these trees tend to occur in clusters so that if you see one tree in a cell then you are likely to see other trees in nearby cells. Why not visit these neighboring cells while in the other?..""

The world is now represented as a graph w/ an adjancy list for each vertex which contains only the neighbors that also have non zero values

For any given source vertex "s", the selected vertex to start an adaptive sample from traverse the graph and find all of the vertices connected to "s"


## Alorithm to obtain this:

<div class="column-left">

Breadth First Search (BFS) on "s" using using the adjacency list generated from the adjlist.adaptive function:
```{}
Algorithm: Breadth First Search (BFS)
- BFS (V,E,s) {
- for each u within V - {s}
    - u.distance <- inf
    - u.parent <- u
- s.d <- 0
- Q <- 0
- Enqueue(Q,s)
- while Q is not empty
    - u <- Dequeue(Q)
    - for each v within G.adj[u]
        - if v.d == inf
        - v.d = u.d + 1
        - v.parent = u
        - Enqueue(Q,v)
- }
```

</div>
<div class="column-right">
- "V" the vertices of the graph

- "E" the adjlist for that graph

    - ie: G.adj[VERTEX] == some adjacency list

- "s" the ID of the vertex to start the BFS on

- "u" some vertex either adjacent to s or s itself, or adjacent to a vertex adjacent to s etc.

The BFS will hit the neighbors of neighbors of "s" which  have non zero values.

</div>

<div class="column-center">
After the BFS any vertex connected to the source vertex "s" will have a distance != infinity
</div>
## BFS ("Adaptive Sampling") in R

```{r, echo = TRUE}
BFS <- function(V,G.adj,s){
  d <- rep(0,nrow(V)*ncol(V))
  p <- rep(0,nrow(V)*ncol(V))
  for(u in 1:length(V)){
    d[u] <- -1
    p[u] <- u
  }
  d[s] <- 0
  Q <- Queue$new()
  Q$push(s)
  id <- c()
  while(!Q$empty()){
    u <- Q$pop()
    id <- c(id,u)
    if(!is.na(G.adj[u])){
      for(v in G.adj[[u]]){
        if(d[v] == -1){
          d[v] <- d[u] + 1
          p[v] <- u
          Q$push(v)
        }
      } 
    }
  }
  return(list("adaptive.sample" = data.frame("y" = sum(V[d != -1]),"m" = length(d[d != -1])),
              "sampled.squares" = data.frame("id" = id, "yi" = V[id]), 
              "d" = matrix(d, nrow = nrow(V), ncol=ncol(V)), 
              "p" = matrix(p, nrow = nrow(V), ncol=ncol(V))
              )
         )
}
```

## What might the distance and parent vectors look like?

$S = 83$
The idea is that after a BFS Traversal of the graph from some node s any visited square will have a distance != $inf$

<div class="column-left">

```{r}
BFS.results <- BFS(figure10.2,adjlist.adaptive(figure10.2),83)
world <- BFS.results$d
plotly.distance <- function(o.world,world,s){
    gg <- melt.matrix(world) %>% 
      mutate(id = getid(x,y,ncol(world)), 
             actual.value = as.vector(o.world),
             distance = value, 
             x = x, 
             y = y) %>% 
      ggplot(aes(x = as.factor(x), 
                 y = as.factor(y),
                 "X" = x,
                 "Y" = y,
                 "id" = id,
                 "distance" = distance,
                 "actual.value" = actual.value)) + 
        geom_raster(aes(fill = value, alpha = value/max(value))) + 
        scale_fill_gradient(low="grey90",high="red") + 
        geom_text(aes(label = ifelse(value == -1,"inf",value))) +
        theme(panel.background = element_blank()) + labs(x = "X",y = "Y") +
        ggtitle(paste("Value of d, distance from s = ",s)) + theme(legend.position = "none")
  
   ggplotly(gg, tooltip = c("X","Y","distance","id","actual.value"))
}
plotly.distance(figure10.2, BFS.results$d,83)
```

All of the visited squares have a value != -1 aka infinity

</div>

<div class="column-right">

```{r}
plotly.parent <- function(o.world,world,name,cols) {
  gg <- melt.matrix(world) %>% 
      mutate(id = getid(x,y,ncol(world)), 
             actual.value = as.vector(o.world), 
             parent = value, x = x, y = y) %>% 
      ggplot(aes(x = as.factor(x), 
                 y = as.factor(y),
                 "X" = x,
                 "Y" = y,
                 "id" = id,
                 "parent" = parent,
                 "actual.value" = actual.value)) + 
        geom_raster(aes(fill = as.factor(value))) + 
        scale_fill_manual(values=cols) +
        geom_text(aes(label = ifelse(value == -1,"inf",value))) +
        theme(panel.background = element_blank()) + labs(x = "X",y = "Y") +
        ggtitle("Value of p, parent of each cell/vertex") + theme(legend.position = "none")
  
   ggplotly(gg, tooltip = c("X","Y","parent","id","actual.value"))
}
cols = rainbow(100, s=.6, v=.9)[sample(1:100,100)]
plotly.parent(figure10.2,BFS.results$p,"Figure 10.2",cols)
```
Most cells are their own parent except for the ones that were visted through BFS Traversal
</div>

## On a lonely node:

$S = 97$

<div class="column-left">

```{r}
BFS.results <- BFS(figure10.2,adjlist.adaptive(figure10.2),97)
plotly.distance(figure10.2,BFS.results$d,97)
```

The value of $-1$ means that there is an "infinite distance" between the two vertices, aka we cannot get there from where we are so these were not sampled for the adaptive sample on "s"

</div>
<div class="column-right">
```{r}
plotly.parent(figure10.2,BFS.results$p,"Figure 10.2",cols)
```
Most cells are their own parent including 97 because BFS did not visit any where
</div>

## What does a sample look like?

<div class="column-left">

The adaptive sample Function:

```{r, echo = TRUE}
adaptive.sample <- function(world, vertecies){
  world.adj <- adjlist.adaptive(world)
  samples <- data.frame("y" = numeric(0),"m" = numeric(0))
  for(i in vertecies){
    samples <- rbind(samples,BFS(world,world.adj,i)$adaptive.sample)
  }
  return(samples)
}

adaptive.sample.figure10.2 <- adaptive.sample(figure10.2, 
                                              c(97,83,56,44,39))
adaptive.sample.figure10.2
```

</div>

<div class="column-right">
```{r}

get.sampled.squares <- function(world,to.sample){
  sampled.squares <- data.frame()
  world.adj <- adjlist.adaptive(world)
  for(i in c(1:length(to.sample))) {
   bfs <- BFS(world,world.adj,to.sample[i]) 
   a <- cbind(bfs$sampled.squares,"BFS" = i)
   sampled.squares <- rbind(sampled.squares,a)
  }
  return(sampled.squares)
}

sampled.squares <- get.sampled.squares(figure10.2, c(97,83,56,44,39))

plotly.sampled <- function(world,sampled.squares,name){
  cols = rainbow(100, s=.6, v=.9)[sample(1:100,100)]
  df <- melt.matrix(world) %>% 
    left_join(sampled.squares, by = "id")
  
  df$BFS[is.na(df$BFS)] <- 0
    
  gg <- df %>%
    ggplot(aes(x = as.factor(x), 
               y = as.factor(y),
               "X" = x,
               "Y" = y,
               "id" = id,
               "value" = value,
               "BFS" = BFS)) + 
      geom_raster(aes(fill = as.factor(BFS), alpha = ifelse(BFS == 0,0, 1))) + 
      # scale_fill_manual(values=cols) +
      geom_text(aes(label = ifelse(BFS == 0 & value == 0, 0, BFS))) +
      theme(panel.background = element_blank()) + labs(x = "X",y = "Y") +
      ggtitle(paste("BFS Groups of",name)) + theme(legend.position = "none")
  
   ggplotly(gg, tooltip = c("X","Y","value","BFS","id"))
}
plotly.sampled(figure10.2,sampled.squares,"Figure 10.2")
```

```{r}
adaptive.sample.confint.fun(adaptive.sample.figure10.2$y,
                            adaptive.sample.figure10.2$m,
                            nrow(adaptive.sample.figure10.2),
                            nrow(figure10.2)*ncol(figure10.2)
                            )[c("mu.hat","v.mu","lower","upper")]

```

</div>

## Each individual BFS:
```{r, echo = TRUE}
bfs.97 <- BFS(figure10.2,figure10.2.adj,97)
bfs.83 <- BFS(figure10.2,figure10.2.adj,83)
bfs.56 <- BFS(figure10.2,figure10.2.adj,56)
bfs.44 <- BFS(figure10.2,figure10.2.adj,44)
bfs.29 <- BFS(figure10.2,figure10.2.adj,29)

bfs.97$adaptive.sample
bfs.97$sampled.squares

bfs.83$adaptive.sample
bfs.83$sampled.squares

bfs.56$adaptive.sample
bfs.56$sampled.squares

bfs.44$adaptive.sample
bfs.44$sampled.squares

bfs.29$adaptive.sample
bfs.29$sampled.squares
```

  
## Adaptive sample on a randomly generated graph:

<div class="column-left">
```{r, echo = TRUE}
set.seed(58039847)
nrow <- 15
ncol <- 30
world.15.30 <- matrix(sample(c(rep(0,25),c(1:10)),
                             nrow*ncol,
                             replace = TRUE),
                nrow,ncol, byrow = TRUE)
```

```{r, fig.align = "center"}
plotly.adjmatrix(world.15.30,"randomly generated 15x30 World w/ seed = 58039847")
```
</div>

<div class="column-right">

```{r, echo = TRUE}
world.15.30.adj <- adjlist.adaptive(world.15.30)
interesting.squares <- c(1,13,53,57,128,227,241,280,358)
adaptive.sample(world.15.30,interesting.squares)
```

```{r,echo = FALSE}
sampled.squares <- get.sampled.squares(world.15.30,interesting.squares)
plotly.sampled(world.15.30,sampled.squares,"world")
```
</div>

## Example on a much larger world
<div class="column-left">
```{r, echo = TRUE}
set.seed(NULL) ## randomize the seed again
nrow <- sample(50:100,1)
ncol <- sample(50:100,1)

world <- matrix(sample(c(rep(0,25),c(1:10)),nrow*ncol,replace = TRUE),
                nrow = nrow,ncol = ncol,byrow=TRUE)
```

```{r}
gg <- world %>%
  melt.matrix() %>% ggplot(aes(x = as.factor(x), y = as.factor(y), label = getid(x,y,ncol), "X" = x, "Y" = y, "ID" = id)) +
    geom_raster(aes(fill = value, alpha = value/max(value))) +
    scale_fill_gradient(low="grey90",high="red") +
    theme(panel.background = element_blank()) + 
    labs(x = "X",y = "Y") + 
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          legend.position = "none") + ggtitle(paste("random",nrow,"x",ncol,"world"))

ggplotly(gg, tooltip = c("X","Y","value","ID"))
```
</div>
<div class="column-right">
```{r, echo = TRUE}
squares.to.sample <- c(sample(which(as.vector(world)!=0),30),
                       sample(which(as.vector(world)==0),30))

adaptive.samples <- adaptive.sample(world, squares.to.sample)

sampled.squares <- get.sampled.squares(world,squares.to.sample)
```

```{r}
cols = rainbow(100, s=.6, v=.9)[sample(1:100,100)]
cols[1] <- "white"

df <- melt.matrix(world) %>% 
  left_join(sampled.squares, by = "id")

df$BFS[is.na(df$BFS)] <- 0

gg <- df %>%
  ggplot(aes(x = as.factor(x), 
             y = as.factor(y),
             "X" = x,
             "Y" = y,
             "id" = id,
             "value" = value,
             "BFS" = BFS)) + 
    geom_raster(aes(fill = as.factor(BFS))) + 
    scale_fill_manual(values=cols) +
    ggtitle(paste("BFS Groups of random",nrow,"x",ncol,"world")) + 
    labs(x = "X",y = "Y") +
    theme(legend.position = "none",
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank())

 ggplotly(gg, tooltip = c("X","Y","value","BFS","id"))
```
</div>
<dic class="column-center">
```{r, echo = TRUE}
adaptive.sample.confint.fun(adaptive.samples$y,
                            adaptive.samples$m,
                            nrow(adaptive.samples),
                            nrow(world)*ncol(world))
```
</div>
## Thank You

- Questions?

ashaffer7@mail.csuchico.edu

Full code: https://github.com/ClassicSours/AdaptiveSampingInR