---
title: "Adaptive Sampling using R"
author: "Aaron Shaffer"
date: "May 3, 2017"
output: slidy_presentation
---
## Adaptive Sampling: (Refresher)
<style>
.column-left{
  float: left;
  width: 50%;
  text-align: left;}
.column-center{
  display: inline-block;
  width: 100%;
  text-align: center;
}
.column-right{
  float: right;
  width: 50%;
  text-align: left;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
pkgs <-c('reshape2','dplyr','ggplot2','plotly','igraph','plotrix')
for(p in pkgs)
  if(p %in% rownames(installed.packages()) == FALSE) {install.packages(p)}
for(p in pkgs)
  suppressPackageStartupMessages(library(p, quietly=TRUE, character.only=TRUE))
rm('p','pkgs')

melt.matrix <- function(world){
  suppressMessages(world %>% 
  data.frame %>% 
  melt() %>% 
  mutate(id = seq_len(nrow(world) * ncol(world)), 
        x = rep(c(1:ncol(world)),nrow(world)),
        y = sort(rep(c(1:nrow(world)),ncol(world)))) %>%
    data.frame() %>% select(value,id,x,y)
  )
}

getid <- function(x,y,ncol){
  return(x + ((y-1) * ncol))
}

plotly.adjmatrix <- function(world, name){
  gg <- melt.matrix(world) %>% 
    ggplot(aes(x = as.factor(x), 
               y = as.factor(y),
               "X" = x,
               "Y" = y,
               "ID" = id)) + 
    geom_raster(aes(fill = value, alpha = value/max(value))) + 
    scale_fill_gradient(low="grey90",high="red") + 
    geom_text(aes(label = ifelse(value,value,""))) +
    theme(panel.background = element_blank(), legend.position = "none") + 
    labs(x = "X",y = "Y") + ggtitle(name)
  
  ggplotly(gg, tooltip = c("X","Y","value","id"))
}

plotly.adjmatrix.id <- function(world, name){
  gg <- melt.matrix(world) %>% 
    ggplot(aes(x = as.factor(x), 
               y = as.factor(y),
               "value" = value,
               "X" = x,
               "Y" = y,
               "ID" = id)) + 
    geom_tile(aes(fill = as.factor(id)),color = "black") + 
    geom_text(aes(label = id)) +
    theme(panel.background = element_blank(), legend.position = "none") + 
    labs(x = "X",y = "Y") + ggtitle(paste("Index values of ",name))
  
  ggplotly(gg, tooltip = c("X","Y","value","id"))
}

```

"10.7: Adaptive Sampling: Suppose that we had a rectangular forest for which you wish to estimate the number of trees of a certain
species. One way to accomplish this is to lay a grid over the forest a sample a certain number of cells. But suppose that these trees tend to occur in clusters so that if you see one tree in a cell then you are likely to see other trees in nearby cells. Why not visit these neighboring cells while in the other? This is an example of adaptive sampling. Instead of a sample size being selected in advance and being fixed the sample size is allowed to vary as new information is coming in."

<div class="column-left">

An unbiased estimator of the mean density per cell in the population is given by:

$\hat{\mu} = \frac{1}{n}\sum\limits_{i=1}^{n}{\frac{y_i}{m_i}} = \frac{1}{n}\sum\limits_{i=1}^{n}{\bar{y_i}}$

Which is the means of the network means.  The estimated variance of this estimator is what you would expect from simple random sampling.

$\hat{V}(\hat{\mu}) = (1 - \frac{n}{N}) \frac{s_\bar{y}^2}{n}$

Where: $\frac{s_\bar{y}^2}{n}$

is the sample variance of the network means

ie: $s_\bar{y}^2 = var(\frac{y_i}{m_i})$

</div>

<div class ="column-right">

In R:

```{r, eval = FALSE, echo = TRUE}
mu.hat <- 1/n * sum(y/m)

s.ybar.2 <- var(y/m)

V.hat.mu.hat <- (1 - n/N) * s.ybar.2/n
```

</div>

## Example from our notes:

<div class="column-left">
$n = 5$

$N = 100$

|     |   |   |   |   |   |
|:---:|:-:|:-:|:-:|:-:|:-:|
|$m_i$| 1 | 4 | 1 | 5 | 2 |
|$y_i$| 0 | 12 | 1 | 15 | 7 |

$\hat{\mu} = \frac{1}{5}\sum{(\frac{0}{1} + \frac{12}{4} + \frac{1}{1} + \frac{15}{5} + \frac{7}{2})} = 2.1$

$s_\bar{y}^2 = var(\frac{y}{m}) = 2.3$

$\hat{V}(\hat{\mu}) = (1 - \frac{5}{100}) \frac{2.3}{5} = 0.437$

</div>
<div class="column-right">
```{r, echo = TRUE}
n <- 5
N <- 100
m <- c(1,4,1,5,2)
y <- c(0,12,1,15,7)
```

```{r, echo = TRUE}
adaptive.sample.confint.fun <- function(y,m,n,N) {
  mu.hat <- 1/n * sum(y/m) 
  s.y.2 <- var(y/m)
  v.mu <- (1 - n/N) * s.y.2/n
  c("mu.hat" = mu.hat, 
    "s.y.2" = s.y.2,
    "v.mu" = v.mu,
    "var" = 2 * sqrt(v.mu),
    "lower" = mu.hat - 2 * sqrt(v.mu),
    "upper" = mu.hat + 2 * sqrt(v.mu))
}

adaptive.sample.confint.fun(y,m,n,N)
```

</div>
<div class="column-center">

So for this sample we estimate that the true mean density of our network was between 0.77 and 3.42 Maple Trees per cell 

</div>

## What might this world have originally looked like?

Data Structure: 2d Matrix

in R: ?matrix

Hint: "lay a grid over the forest"

<div class="column-left">
```{r, echo = TRUE}
example.world <- matrix(c( 1, 0, 3, 3, 0, 0,-1,
                           0, 0, 3, 0, 3, 0,-1,
                           3, 0, 0, 0, 0, 0, 0,
                           0, 4, 0, 0, 0, 3, 3,
                           0, 0, 0, 3, 3, 3, 0,
                          -1,-1, 0, 0, 0, 0, 0, 
                          -1,-1,-1,-1,-1,-1,-1),
                        nrow = 7, ncol = 7,byrow=TRUE)
example.world
```


If you want to rebuild a world given some adaptive samples:

Think of a world like a game of minesweeper.

Since we know the y and m for a few samples we know a couple of things.  Anything around these samples has to have a value of 0, as defined by what it means to take an adaptive sample.

Any square in the representation of the world that doesnt touch a known sample has some unknown value "-1"

</div>
<div class="column-right">
```{r}
plotly.adjmatrix(example.world,"example.world, -1 means you cannot infer what value might be there")
```
</div>

## How do we navigate this world on the computer?

<div class="column-left">

First lets look at another example world:

```{r, echo = TRUE}
figure10.2 <- matrix(rep(rep(0,10),10),
  nrow = 10, ncol = 10)

figure10.2[,10] <- c(4,0,0,0,0,0,0,0,0,7)
figure10.2[,9]  <- c(3,0,2,1,0,0,0,0,4,1)
figure10.2[,8]  <- c(0,0,0,4,5,0,0,3,2,0)
figure10.2[,7]  <- c(5,4,0,0,0,0,0,0,0,0)
figure10.2[,6]  <- c(0,0,0,2,0,1,0,0,7,0)
figure10.2[,5]  <- c(0,5,0,4,0,0,0,0,0,0)
figure10.2[,4]  <- c(0,0,3,0,0,4,0,3,0,0)
figure10.2[,3]  <- c(0,1,0,0,0,3,0,0,4,0)
figure10.2[,2]  <- c(0,0,0,2,6,0,0,0,0,0)
figure10.2[,1]  <- c(0,4,0,4,0,0,2,3,0,0)
```

In order to navigate this worlds we first need to know a couple of definitions

</div>

<div class="column-right">

```{r}
plotly.adjmatrix(figure10.2,"Figure 10.2")
```
</div>

## Definition 1: "Adjacency List"

"In graph theory and computer science and adjacency list is a collection of unordered lists used to represent a finite graph.  Each list represents the set of neighbors of a vertex of a graph"

1. Represent the matrix as a graph where each cell of the matrix is the vertex on a graph.

Three functions created 

Adjlist function -> give it a world as a matrix object, gives you back and adjacency list

getEdges function -> give it a world as a matrix object, returns an adjacency list that the igraph package can use to create a graph

Adjlist.adaptive function -> given a world as a matrix object returns the adjacency list for a world that obeys the rules of an adaptive sampling grid

## How to generate the Adjlist:

<div class="column-left">

Core idea:

"column" up and down

"row" left and right

"%%" modulo, finds the remainder in a division

```{}
for each index "i" of the matrix:
    - Is "i" in the Top Row?
          "is top" - i > (nrow * ncol) - ncol
    - Is "i" in the Bottom Row?
          "is bottom" - i <= ncol
    - Is "i" in the Leftmost column?
          "is left" - (i - 1) %% ncol == 0
    - Is "i" in the Rightmost Column?
          - i %% ncol == 0
          
    if (I'm not Leftmost column)
    left <- i - 1
    
    if (I'm not in the Leftmost column or Topmost row)
    upleft <- i - 1 + ncol
    
    if (I'm not in the Topmost Row)
    up <- i + ncol
    
    if (I'm not in the Topmost row or Rightmost column)
    upright <- i + 1 + ncol
    
    if (I'm not in the Rightmost column)
    right <- i + 1
    
    if (I'm not in the Bottommost row or the Rightmost column)
    downright <- i + 1 - ncol
    
    if (I'm not in the Bottommost row)
    down <- i - ncol
```
</div>

<div class="column-right">

plotly!, mouse over graphs for more information!

```{r}
sprintf("ncol: %d",ncol(figure10.2))
sprintf("nrow: %d",nrow(figure10.2))
plotly.adjmatrix.id(figure10.2, "Figure 10.2")
```
</div>

## Adjlist Functions:

<div class="column-left">
```{r, echo = TRUE}
adjlist <- function(world){
  nrow <- nrow(world)
  ncol <- ncol(world)
  
  G <- as.vector(world)
  
  adj.list <- rep(list(c()),nrow*ncol)  
  for(index in 1:length(G)){
    last.column <- index %% ncol == 0
    first.column <- (index - 1) %% ncol == 0
    first.row <- index <= ncol
    last.row <- index > (nrow * ncol - ncol)
    
    left <- ifelse(first.column,NA,index - 1)
    up.left <- ifelse(last.row | first.column,NA,index + ncol - 1)
    up <- ifelse(last.row,NA,index + ncol)
    up.right <- ifelse(last.row | last.column, NA,index + ncol + 1)
    right <- ifelse(last.column, NA, index + 1)
    down.right <- ifelse(first.row | last.column, NA, index - ncol + 1)
    down <- ifelse(first.row, NA, index - ncol)
    down.left <- ifelse(first.row | first.column , NA, index - ncol - 1)
    
    adjlist.i <- c()
    if(!is.na(left))
      adjlist.i <- c(adjlist.i,"W" = left)
    if(!is.na(up.left))
      adjlist.i <- c(adjlist.i,"NW" = up.left)
    if(!is.na(up))
      adjlist.i <- c(adjlist.i,"N" = up)
    if(!is.na(up.right))
      adjlist.i <- c(adjlist.i,"NE" = up.right)
    if(!is.na(right))
      adjlist.i <- c(adjlist.i,"E"= right)
    if(!is.na(down.left))
      adjlist.i <- c(adjlist.i,"SE" = down.left)
    if(!is.na(down))
      adjlist.i <- c(adjlist.i,"S" = down)
    if(!is.na(down.right))
      adjlist.i <- c(adjlist.i,"SW" = down.right)
    
    adj.list[[index]] <- adjlist.i
  }
  return(adj.list)
}
```
</div>

<div class="column-right">
```{r, echo = TRUE}
adjlist.adaptive <- function(world){
  nrow <- nrow(world)
  ncol <- ncol(world)
  
  G <- as.vector(world)
  
  adj.list <- rep(list(c()),nrow*ncol)  
  for(index in 1:length(G)){
    if(G[index] != 0) {
      last.column <- index %% ncol == 0
      first.column <- (index - 1) %% ncol == 0
      first.row <- index <= ncol
      last.row <- index > (nrow * ncol - ncol)
      
      left <- ifelse(first.column,NA,index - 1)
      up.left <- ifelse(last.row | first.column,NA,index + ncol - 1)
      up <- ifelse(last.row,NA,index + ncol)
      up.right <- ifelse(last.row | last.column, NA,index + ncol + 1)
      right <- ifelse(last.column, NA, index + 1)
      down.right <- ifelse(first.row | last.column, NA, index - ncol + 1)
      down <- ifelse(first.row, NA, index - ncol)
      down.left <- ifelse(first.row | first.column , NA, index - ncol - 1)
      
      to.return <- c()
      if(!is.na(left)) {
        if(G[left] != 0) {
          to.return <- c(to.return,"W" = left) 
        }
      }
      if(!is.na(up.left)){
        if(G[up.left] != 0) {
          to.return <- c(to.return,"NW" = up.left)  
        }
      }
      if(!is.na(up)) {
        if(G[up] != 0){
          to.return <- c(to.return,"N" = up)
        }
      }
      if(!is.na(up.right)) {
        if(G[up.right] != 0){
          to.return <- c(to.return,"NE" = up.right)
        }
      }
      if(!is.na(right)) {
        if(G[right] != 0){
          to.return <- c(to.return,"E"= right)
        }
      }
      if(!is.na(down.left)) {
        if(G[down.left] != 0){
          to.return <- c(to.return,"SE" = down.left)
        }
      }
      if(!is.na(down)) {
        if(G[down] != 0){
          to.return <- c(to.return,"S" = down)
        }
      }
      if(!is.na(down.right)) {
        if(G[down.right] != 0){
          to.return <- c(to.return,"SW" = down.right)
        }
      }
      
      adj.list[[index]] <- to.return
    } else {
      adj.list[[index]] <- c("NA" = NA)
    }
  }
  return(adj.list)
}
```
</div>

## Non adaptive figure 10.2

<div class="column-left">
```{r}
plotly.adjmatrix.id(figure10.2, "Figure 10.2")
```

```{r, echo =  TRUE}
figure10.2.adjlist <- adjlist(figure10.2)
figure10.2.adjlist
```

</div>

```{r}
getEdges <- function(world,adaptive){
  if(adaptive == TRUE)
    world.adj <- adjlist.adaptive(figure10.2)
  else
    world.adj <- adjlist(figure10.2)
  edge.list <- c()
  for(i in 1:length(world.adj)){
    for(j in 1:length(world.adj[[i]])){
      if(i <= ifelse(!gtools::invalid(world.adj[[i]][j]),world.adj[[i]][j],0)){
        edge.list <- c(edge.list,i,ifelse(!gtools::invalid(world.adj[[i]][j]),world.adj[[i]][j],i))
      }
    }
  }
  edge.list
}

```

<div class="column-right">

```{r, echo = TRUE}
g1 <- graph(edges = rev(getEdges(figure10.2,FALSE)), 
            directed = F, n = 100)
plot(g1, layout=layout_on_grid)
```

</div>


## Adj list for a function following the rules of adaptive sampling

<div class ="column-left">

```{r}
plotly.adjmatrix(figure10.2, "Figure 10.2")
```

```{r, echo = TRUE}
figure10.2.adj <- adjlist.adaptive(figure10.2)
figure10.2.adj
```

</div>
<div class="column-right">
```{r, echo = TRUE}
g2 <- graph(edges = rev(getEdges(figure10.2,TRUE)), 
            directed = F, n = 100)
plot(g2, layout=layout_on_grid)
```
</div>

## Compare the differences in the adjacency list

<div class="column-left">
```{r,echo = TRUE}
g1 <- graph(edges = rev(getEdges(figure10.2,FALSE)), 
            directed = F, n = 100)
plot(g1, layout=layout_on_grid)
```

```{r}
plotly.adjmatrix.id(figure10.2,"Figure 10.2")
```

```{r}
adjlist(figure10.2)
```

</div>

<div class="column-right">
```{r, echo = TRUE}
g2 <- graph(edges = rev(getEdges(figure10.2,TRUE)), 
            directed = F, n = 100)
plot(g2, layout=layout_on_grid)
```

```{r}
plotly.adjmatrix(figure10.2,"Figure 10.2")
```

```{r}
adjlist.adaptive(figure10.2)
```
</div>



## Definition 2: "Queue"

A Queue is essentially a fancy vector

Add things to the back of the Queue ("push")

Remove things from the front of the Queue ("dequeue")

```{r, echo = TRUE}
Queue <- setRefClass(Class = "Queue",
                     fields = list(
                       name = "character",
                       data = "list"
                     ),
                     methods = list(
                       size = function() {
                         'Returns the number of items in the queue.'
                         return(length(data))
                       },
                       #
                       push = function(item) {
                         'Inserts element at back of the queue.'
                         data[[size()+1]] <<- item
                       },
                       #
                       pop = function() {
                         'Removes and returns head of queue (or raises error if queue is empty).'
                         if (size() == 0) stop("queue is empty!")
                         value <- data[[1]]
                         data[[1]] <<- NULL
                         value
                       },
                       #
                       initialize=function(...) {
                         callSuper(...)
                         #
                         # Initialise fields here (place holder)...
                         #
                         .self
                       },
                       empty=function(...){
                         return(size() == 0)
                       }
                     )
)
```

## How do we actually take the sample?

Graph theory & computer science to the rescue!

```{}
Algorithm: Breadth First Search (BFS)
- BFS (V,E,s) {
- for each u within V - {s}
    - u.distance <- inf
    - u.parent <- u
- s.d <- 0
- Q <- 0
- Enqueue(Q,s)
- while Q is not empty
    - u <- Dequeue(Q)
    - for each v within G.adj[u]
        - if v.d == inf
        - v.d = u.d + 1
        - v.parent = u
        - Enqueue(Q,v)
- }
```

- "V" the vertices of the graph

- "E" the adjlist for that graph

    - ie: G.adj[VERTEX] == some adjacency list

- "s" the ID of the vertex to start the BFS on

- "u" some vertex either adjacent to s or s itself, or adjacent to a vertex adjacent to s etc.

## BFS ("Adaptive Sampling")

```{r, echo = TRUE}
BFS <- function(V,G.adj,s){
  d <- rep(0,nrow(V)*ncol(V))
  p <- rep(0,nrow(V)*ncol(V))
  for(u in 1:length(V)){
    d[u] <- -1
    p[u] <- u
  }
  d[s] <- 0
  Q <- Queue$new()
  Q$push(s)
  id <- c()
  while(!Q$empty()){
    u <- Q$pop()
    id <- c(id,u)
    if(!is.na(G.adj[u])){
      for(v in G.adj[[u]]){
        if(d[v] == -1){
          d[v] <- d[u] + 1
          p[v] <- u
          Q$push(v)
        }
      } 
    }
  }
  return(list("adaptive.sample" = data.frame("y" = sum(V[d != -1]),"m" = length(d[d != -1])),
              "sampled.squares" = data.frame("id" = id, "yi" = V[id]), 
              "d" = matrix(d, nrow = nrow(V), ncol=ncol(V)), 
              "p" = matrix(p, nrow = nrow(V), ncol=ncol(V))))
}
```

## What might the distance and parent vectors look like?

$S = 83$
The idea is that after a BFS Traversal of the graph from some node s any visited square will have a distance != $inf$

<div class="column-left">

```{r}
BFS.results <- BFS(figure10.2,adjlist.adaptive(figure10.2),83)
world <- BFS.results$d
plotly.distance <- function(o.world,world,s){
    gg <- melt.matrix(world) %>% 
      mutate(id = getid(x,y,ncol(world)), 
             actual.value = as.vector(o.world),
             distance = value, 
             x = x, 
             y = y) %>% 
      ggplot(aes(x = as.factor(x), 
                 y = as.factor(y),
                 "X" = x,
                 "Y" = y,
                 "id" = id,
                 "distance" = distance,
                 "actual.value" = actual.value)) + 
        geom_raster(aes(fill = value, alpha = value/max(value))) + 
        scale_fill_gradient(low="grey90",high="red") + 
        geom_text(aes(label = ifelse(value == -1,"inf",value))) +
        theme(panel.background = element_blank()) + labs(x = "X",y = "Y") +
        ggtitle(paste("Value of d, distance from s = ",s)) + theme(legend.position = "none")
  
   ggplotly(gg, tooltip = c("X","Y","distance","id","actual.value"))
}
plotly.distance(figure10.2, BFS.results$d,83)
```

All of the visited squares have a value != -1 aka infinity

</div>

<div class="column-right">

```{r}
plotly.parent <- function(o.world,world,name,cols) {
  gg <- melt.matrix(world) %>% 
      mutate(id = getid(x,y,ncol(world)), 
             actual.value = as.vector(o.world), 
             parent = value, x = x, y = y) %>% 
      ggplot(aes(x = as.factor(x), 
                 y = as.factor(y),
                 "X" = x,
                 "Y" = y,
                 "id" = id,
                 "parent" = parent,
                 "actual.value" = actual.value)) + 
        geom_raster(aes(fill = as.factor(value))) + 
        scale_fill_manual(values=cols) +
        geom_text(aes(label = ifelse(value == -1,"inf",value))) +
        theme(panel.background = element_blank()) + labs(x = "X",y = "Y") +
        ggtitle("Value of p, parent of each cell/vertex") + theme(legend.position = "none")
  
   ggplotly(gg, tooltip = c("X","Y","parent","id","actual.value"))
}
cols = rainbow(100, s=.6, v=.9)[sample(1:100,100)]
plotly.parent(figure10.2,BFS.results$p,"Figure 10.2",cols)
```
Most cells are their own parent except for the ones that were visted through BFS Traversal
</div>

## On a lonely node:

$s = 97$

<div class="column-left">

```{r}
BFS.results <- BFS(figure10.2,adjlist.adaptive(figure10.2),97)
plotly.distance(figure10.2,BFS.results$d,97)
```

All of the visited squares have a value != -1 (infinity)

The value of $-1$ means that there is an "infinite distance" between the two vertices, aka we cannot get there from where we are

</div>
<div class="column-right">
```{r}
plotly.parent(figure10.2,BFS.results$p,"Figure 10.2",cols)
```
Most cells are their own parent including 97 because BFS did not visit any where
</div>

## What does a sample look like?

<div class="column-left">

```{r}
plotly.adjmatrix(figure10.2, "Figure 10.2")
```

```{r, echo = TRUE}
adaptive.sample <- function(world, vertecies){
  world.adj <- adjlist.adaptive(world)
  samples <- data.frame("y" = numeric(0),"m" = numeric(0))
  for(i in vertecies){
    samples <- rbind(samples,BFS(world,world.adj,i)$adaptive.sample)
  }
  return(samples)
}
adaptive.sample.figure10.2 <- adaptive.sample(figure10.2, 
                                              c(97,83,56,44,39))
adaptive.sample.figure10.2

adaptive.sample.confint.fun(adaptive.sample.figure10.2$y,
                            adaptive.sample.figure10.2$m,
                            nrow(adaptive.sample.figure10.2),
                            nrow(figure10.2)*ncol(figure10.2)
                            )[c("mu.hat","v.mu","lower","upper")]
```
</div>

<div class="column-right">

Each individual BFS:

```{r, echo = TRUE}
bfs.97 <- BFS(figure10.2,figure10.2.adj,97)
bfs.97$adaptive.sample
bfs.97$sampled.squares

bfs.83 <- BFS(figure10.2,figure10.2.adj,83)
bfs.83$adaptive.sample
bfs.83$sampled.squares

bfs.56 <- BFS(figure10.2,figure10.2.adj,56)
bfs.56$adaptive.sample
bfs.56$sampled.squares

bfs.44 <- BFS(figure10.2,figure10.2.adj,44)
bfs.44$adaptive.sample
bfs.44$sampled.squares

bfs.29 <- BFS(figure10.2,figure10.2.adj,29)
bfs.29$adaptive.sample
bfs.29$sampled.squares
```
</div>

## The sampled squares:

<div class="column-left">
```{r}
sampled.squares <- rbind(
  cbind(bfs.97$sampled.squares,"BFS" = 1),
  cbind(bfs.83$sampled.squares,"BFS" = 2),
  cbind(bfs.56$sampled.squares,"BFS" = 3),
  cbind(bfs.44$sampled.squares,"BFS" = 4),
  cbind(bfs.29$sampled.squares,"BFS" = 5))
sampled.squares
```
</div>
<div class = "column-right">
```{r}
figure10.2.sampled.squares <- rbind(
  cbind(bfs.97$sampled.squares,"BFS" = 1),
  cbind(bfs.83$sampled.squares,"BFS" = 2),
  cbind(bfs.56$sampled.squares,"BFS" = 3),
  cbind(bfs.44$sampled.squares,"BFS" = 4),
  cbind(bfs.29$sampled.squares,"BFS" = 5))

plotly.sampled <- function(world,sampled.squares,name){
  cols = rainbow(100, s=.6, v=.9)[sample(1:100,100)]
  df <- melt.matrix(world) %>% 
    left_join(sampled.squares, by = "id")
  
  df$BFS[is.na(df$BFS)] <- 0
    
  gg <- df %>%
    ggplot(aes(x = as.factor(x), 
               y = as.factor(y),
               "X" = x,
               "Y" = y,
               "id" = id,
               "value" = value,
               "BFS" = BFS)) + 
      geom_raster(aes(fill = as.factor(BFS), alpha = ifelse(BFS == 0,0, 1))) + 
      # scale_fill_manual(values=cols) +
      geom_text(aes(label = ifelse(BFS == 0 & value == 0, 0, BFS))) +
      theme(panel.background = element_blank()) + labs(x = "X",y = "Y") +
      ggtitle(paste("BFS Groups of",name)) + theme(legend.position = "none")
  
   ggplotly(gg, tooltip = c("X","Y","value","BFS","id"))
}
plotly.sampled(figure10.2,sampled.squares,"Figure 10.2")
```
</div>

## Adaptive sample on some random graph:

<div class="column-left">
```{r, echo = TRUE}
set.seed(58039847)
nrow <- 15
ncol <- 30
world.15.30 <- matrix(sample(c(rep(0,25),c(1:10)),
                             nrow*ncol,
                             replace = TRUE),
                nrow,ncol, byrow = TRUE)
```

```{r, fig.align = "center"}
plotly.adjmatrix(world.15.30,"randomly generated 15x30 World w/ seed = 58039847")
```
</div>

<div class="column-right">

```{r, echo = TRUE}
world.15.30.adj <- adjlist.adaptive(world.15.30)
interesting.squares <- c(1,13,53,57,128,227,241,280,358)
adaptive.sample(world.15.30,interesting.squares)
```

```{r,echo = FALSE}
get.sampled.squares <- function(world,to.sample){
  sampled.squares <- data.frame()
  world.adj <- adjlist.adaptive(world)
  for(i in c(1:length(to.sample))) {
   bfs <- BFS(world,world.adj,to.sample[i]) 
   a <- cbind(bfs$sampled.squares,"BFS" = i)
   sampled.squares <- rbind(sampled.squares,a)
  }
  return(sampled.squares)
}
sampled.squares <- get.sampled.squares(world.15.30,interesting.squares)
plotly.sampled(world.15.30,sampled.squares,"world")
```
</div>

## lets get really random
<div class="column-left">
```{r, echo = TRUE}
set.seed(NULL) ## randomize the seed again
nrow <- sample(50:100,1)
ncol <- sample(50:100,1)
world <- matrix(sample(c(rep(0,25),c(1:10)),nrow*ncol,replace = TRUE),
                nrow = nrow,ncol = ncol,byrow=TRUE)
```

```{r}
gg <- world %>%
  melt.matrix() %>% ggplot(aes(x = as.factor(x), y = as.factor(y), label = getid(x,y,ncol), "X" = x, "Y" = y, "ID" = id)) +
    geom_raster(aes(fill = value, alpha = value/max(value))) +
    scale_fill_gradient(low="grey90",high="red") +
    theme(panel.background = element_blank()) + 
    labs(x = "X",y = "Y") + 
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          legend.position = "none") + ggtitle(paste("random",nrow,"x",ncol,"world"))

ggplotly(gg, tooltip = c("X","Y","value","ID"))
```
</div>
<div class="column-right">
```{r, echo = TRUE}
squares.to.sample <- c(sample(which(as.vector(world)!=0),30),
                       sample(which(as.vector(world)==0),30))

adaptive.samples <- adaptive.sample(world, squares.to.sample)

sampled.squares <- get.sampled.squares(world,squares.to.sample)
```

```{r}
cols = rainbow(100, s=.6, v=.9)[sample(1:100,100)]
cols[1] <- "white"

df <- melt.matrix(world) %>% 
  left_join(sampled.squares, by = "id")

df$BFS[is.na(df$BFS)] <- 0

gg <- df %>%
  ggplot(aes(x = as.factor(x), 
             y = as.factor(y),
             "X" = x,
             "Y" = y,
             "id" = id,
             "value" = value,
             "BFS" = BFS)) + 
    geom_raster(aes(fill = as.factor(BFS))) + 
    scale_fill_manual(values=cols) +
    ggtitle(paste("BFS Groups of random",nrow,"x",ncol,"world")) + 
    labs(x = "X",y = "Y") +
    theme(legend.position = "none",
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank())

 ggplotly(gg, tooltip = c("X","Y","value","BFS","id"))
```
</div>
<dic class="column-center">
```{r, echo = TRUE}
adaptive.sample.confint.fun(adaptive.samples$y,
                            adaptive.samples$m,
                            nrow(adaptive.samples),
                            nrow(world)*ncol(world))
```
</div>
## Thank You

- Questions?

ashaffer7@mail.csuchico.edu